---
title: "Turkey Functions"
author: "Austin Acree; Henry Traynor; Mike Nolan-Tamariz"
date: "2024-11-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Collections of functions implemented throughout this project


#### Henry's intervalAnalysis
Developed to different statistics for data smoothing

```{r}
intervalAnalysis <- function(df.sample, fun.call, dataCol, windowLength, windowStep) {
  colNum = which(colnames(df.sample)==dataCol)
  #calculation of statistic values
  stat <- rollapply(df.sample[,colNum],
                    FUN=fun.call,
                    width=windowLength,
                    by = windowStep,
                    by.column = TRUE,
                    align = "right")
  
  timestamp <- seq(from=df.sample$dt[windowLength], to=df.sample$dt[nrow(df.sample)], length.out=length(stat))
  #combining stat values with time indices
  df.stat <- data.frame(timestamp,stat)
  
  return(df.stat)
}

```

#### shrink
Subset smaller data ranges; copy paste into code to set date range
Applied to raw MoveBank csv

```{r}
shrink <- function (data) {
  start_date <- ymd_hms("2024-02-26 05:16:00")
  end_date <- ymd_hms("2024-05-15 12:00:00")
  
  data <- subset(data, timestamp >= start_date & timestamp <= end_date)
  return(x)}
```

#### retime
Sets new dates and times for processing through the given code. Sets time/date as POSITXct items. 

```{r}
retime <- function(data) {
  data$Month = format(as.POSIXct(data$timestamp), '%m')
  data$Day = format(as.POSIXct(data$timestamp), '%d')
  data$Year = format(as.POSIXct(data$timestamp), '%Y')
  data$Date = format(as.POSIXct(data$timestamp), '%Y-%m-%d')
  data$Time = format(as.POSIXct(data$timestamp), '%H:%M:%S')
  return(data)}

```

#### reName
Sets simplified columns name (from raw MoveBank csv). Needed for original code string. 
Capitalize 'N' to distinguish from base dplyr function
```{r}
reName <- function (dataframe) {
  dataframe <-dataframe %>% 
    rename(ID="individual-local-identifier", accel=`eobs:accelerations-raw`)
  return(dataframe)
}
```

#### expandAccel
To separate acceleration values. Pulled from original code.

```{r}
expandAccel <- function (data) {
  
  # Function to split 'accel' column into groups of three values (x, y, z axes)
  split_accel_data <- function(accel_col) {
    accel_values <- unlist(strsplit(accel_col, " "))
    matrix(accel_values, ncol = 3, byrow = TRUE)
  }
  
  # Initialize an empty dataframe to store the expanded data
  expanded_data <- data.frame()
  
  # Iterate through each row in the dataframe
  for (i in 1:nrow(data)) {
    # Split accel column into x, y, z values
    accel_matrix <- split_accel_data(data$accel[i])
    
    # Create a temporary dataframe for the current row, repeating the timestamp, date, year, and tag-local-identifier for each sample
    temp_data <- data.frame(
      Time = rep(data$Time[i], nrow(accel_matrix)),
      Date = rep(data$Date[i], nrow(accel_matrix)),
      Year = rep(data$Year[i], nrow(accel_matrix)),
      ID = rep(data$ID[i], nrow(accel_matrix)),
      sample_num = 1:nrow(accel_matrix),  # Add a sample number for each sample
      x_axis = accel_matrix[, 1],
      y_axis = accel_matrix[, 2],
      z_axis = accel_matrix[, 3]
    )
    
    # Append the temporary dataframe to the expanded dataframe
    expanded_data <- rbind(expanded_data, temp_data)
  }
  expanded_data$x_axis=as.numeric(expanded_data$x_axis)
  expanded_data$y_axis=as.numeric(expanded_data$y_axis)
  expanded_data$z_axis=as.numeric(expanded_data$z_axis)
  
  
  return(expanded_data)
}
```

#### simplify
Pulled from Henry's code to simplify the datframe. I have not, however troubleshot this yet. 

```{r}
simplify <- function (dataframe) {
  smaller.df <- dataframe %>%
  select("ID", "dt", "x_avg", "y_avg", "z_avg")
  return(smaller.df)
}
```

#### condenseBurst
Takes the average of values on an accel axis per burst.
This could probably be expanded to apply across all three axes at once, but has to be applied individually for now. 

```{r}
condense_burst <- function(column, window_size = 40) {
  sapply(seq_along(column), function(i) {
    start <- floor((i - 1) / window_size) * window_size + 1
    end <- min(start + window_size - 1, length(column))
    mean(column[start:end])
  })
}
```

#### create "dt" column
This could probably get integrated into the "retime" function, so long as it does not get lost in the expandAccel function later. Could also get worked into another function such as "simplify".

```{r}
create.dt <- function (df) { 
  df$DateTime = paste(df$Date, df$Time)
df$dt = ymd_hms(df$DateTime, tz="GMT")
df$dt = df$dt - hours(5)
return(df)
}

createdt <- function (dataframe) {
dataframe$DateTime = paste(dataframe$Date, dataframe$Time)
dataframe$DT = ymd_hms(dataframe$DateTime,tz="GMT")
}
```

